Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}

### 2017-08-14 ms.tex
conclu

Based on these assumptions, we fit for envelope metallicity and core mass using an MCMC with uniform priors on both parameters.  The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples, which we thinned to $10^5$.  The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was 18 (the value for water), using the approach of \cite{fortney13}. Figure\,\ref{fig:metal_prior} shows the results.  We find that the inferred atmospheric metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius). Larger core masses and envelope metallicities are allowed for the younger stellar age since planets cool and contract as they age.  For the younger (older) stellar age, we find an upper limit on the core mass of FIXME and an upper limit on atmospheric metallicity of FIXME.

@@

We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}. With the most efficient haze production and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations. However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.  %Sulfur-based hazes are another possibility \citep{gao17}.
Further modeling of the condensates is needed for a satisfactory explanation of the data.

%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.  
\section{Conclusions} \label{sec:discuss}
We analyzed the atmospheric composition of WASP-107b based on its measured near-infrared transmission spectrum and interior structure models of the planet's mass and radius.  Key results from this analysis include:

\begin{itemize}
\item{The upper limit on atmospheric metallicity is $30\times$ solar (assuming the star is young) or FIXME solar for an older system.  Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.  Based on the Solar System trend of increasing planet metallicity with decreasing planet mass, WASP-107b is expected to have a metallicity of $30\times$ solar, just at the edge of our allowed range \citep{kreidberg14b}. Further characterization of the stellar age is needed to determine whether WASP-107b is consistent with expectations, or discrepant with the Solar System trend, which would suggest an intriguing difference in its formation history. Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{morley17}. Together, these exoplanets suggest a greater diversity of compositions than exists in the Solar System.}
\item{The methane abundance is depleted relative to expectations for a solar abundance pattern. This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.  Another possibility is that the planet has a hot interior ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}. Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}.} 
\item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).  Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed . Clouds form too deep in the atmosphere . \citep{crossfield17}} GJ 1214b
\end{itemize}

>>>>

Based on these assumptions, we fit for envelope metallicity and core mass using an MCMC with uniform priors on both parameters.  The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples, which we thinned to $10^5$.  The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was 18 (the value for water), using the approach of \cite{fortney13}. Figure\,\ref{fig:metal_prior} shows the results.  We find an upper limit on the metallicity of $30\times$ solar for the young stellar age range.  Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.  For the older age, the upper limit is even lower (FIXME), because planets cool and contract as they age \citep{fortney08}.  The inferred metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius). The upper limit on the core mass is FIXME.

@@

We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}. With the most efficient haze production and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations. However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.  Further modeling of the condensates is needed for a satisfactory explanation of the data.

%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.  
\section{Discussion and Conclusions} \label{sec:discuss}
We analyzed the atmospheric composition of WASP-107b based on retrieval of its near-infrared transmission spectrum and interior structure models of the planet's mass and radius.  Key results from this analysis include:

\begin{itemize}
\item{\emph{The upper limit on atmospheric metallicity is $30\times$ solar.} Our estimated upper limit is at the edge of consistency with the Solar System metal enrichment pattern, with predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.  If the star is old (FIXME Gyr), the upper limit on metallicity would be even lower, implying that WASP-107b has a different formation history from the Solar System planets. Further characterization of the stellar age is needed to explore this scenario.  Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{morley17}. Together, these exoplanets hint at a greater diversity of compositions than exists in the Solar System.}
\item{The methane abundance is depleted relative to expectations for a solar abundance pattern. This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.  Another possibility is that the planet has a hot interior ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}. Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}. Further observations of the transmission spectrum over a broader wavelength range will refine the C/O estimate and help differentiate between these two scenarios.} 
\item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).  Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observations. The atmospheric scale height is so large (800 km) that clouds form too deep in the atmosphere to loft to the needed height. Hydrocarbon hazes that form at high altitudes can match the data, but these require methane photolysis as a precursor, and methane is depleted in the atmosphere.  Other haze precursors such as $H_2S$ are a possibility and would be worth exploring in future analyses \citep{gao17}.
\citep{crossfield17}} GJ 1214b
\end{itemize}
